---
title: "fused_matrix_visualisation"
author: "Hani Jieun Kim"
date: "11/03/2020"
output: 
  BiocStyle::html_document:
    code_folding: hide
    number_sections: yes
    toc: yes  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Fused matrix visualisation

The script reproduces the UMAP visualisation of the fused matrix in Figure 2c-e in the main manuscript. It is shown in Figures \@ref(fig:plot_full), \@ref(fig:plot_modality), and \@ref(fig:plot_markers).
```{r load, echo=FALSE}
suppressMessages(library(uwot))
suppressMessages(library(ggplot2))
suppressMessages(library(ggthemes))
suppressMessages(library(tidyverse))
suppressMessages(library(preprocessCore))
suppressMessages(library(SingleCellExperiment))

load("fusedMatrix_CiteFuse.RData", verbose=T)
load("cluster_labels.RData", verbose = T)
load("annotation_colors.RData", verbose=T)
load("CiteFuse_filtered.RData", verbose=T)
```

Generate UMAP through the ```uwot``` package.
```{r}
set.seed(2019)
SNF_umap <- uwot::umap(fusedMatrix_CiteFuse)

dftoPlot <- 
    data.frame(
        UMAP1 = SNF_umap[, 1],
        UMAP2 = SNF_umap[, 2]
    )
```

# Plot UMAP of the full dataset
```{r plot_full, fig.cap="UMAP of the full human PBMC CITE-seq dataset."}
gg.full <- 
    ggplot(dftoPlot, aes(x = UMAP1, y = UMAP2, color = as.factor(CiteFuse_cluster_labels))) +
    geom_point(alpha = 0.5) +
    scale_color_manual(values = annotation_colors$W_colors) +
    theme_bw() +
    theme(aspect.ratio = 1, legend.position = "bottom") 

gg.full

gg.full %>% ggsave(filename="Figure2c.pdf", width=5, height=5.5)
```


# Plot UMAP of CD4+ T-cells highlighted by clustering output from different data modalities
```{r plot_modality, fig.cap="UMAP of CD4+ T-cells highlighted by clusters."}
gg.CD4_Tcells.CiteFuse <- 
    ggplot(dftoPlot, aes(x = UMAP1, y = UMAP2, color = as.factor(CiteFuse_cluster_labels))) +
    geom_point(alpha = 0.5) +
    coord_cartesian(xlim=c(8,15), ylim=c(-7,3)) + 
    scale_color_manual(values = annotation_colors$W_colors) +
    theme_bw() +
    ggtitle("CiteFuse") + 
    theme(aspect.ratio = 1, legend.position = "none") 

gg.CD4_Tcells.ADT <- 
    ggplot(dftoPlot, aes(x = UMAP1, y = UMAP2, color = as.factor(ADT_cluster_labels))) +
    geom_point(alpha = 0.5) +
    coord_cartesian(xlim=c(8,15), ylim=c(-7,3)) + 
    scale_color_manual(values = annotation_colors$W1_colors) +
    theme_bw() +
    ggtitle("ADT alone") + 
    theme(aspect.ratio = 1, legend.position = "none") 

gg.CD4_Tcells.RNA <- 
    ggplot(dftoPlot, aes(x = UMAP1, y = UMAP2, color = as.factor(RNA_cluster_labels))) +
    geom_point(alpha = 0.5) +
    coord_cartesian(xlim=c(8,15), ylim=c(-7,3)) + 
    scale_color_manual(values = annotation_colors$W2_colors) +
    theme_bw() +
    ggtitle("RNA alone") + 
    theme(aspect.ratio = 1, legend.position = "none") 

ggpubr::ggarrange(gg.CD4_Tcells.CiteFuse, 
                  gg.CD4_Tcells.ADT, 
                  gg.CD4_Tcells.RNA, ncol=3) 

ggpubr::ggarrange(gg.CD4_Tcells.CiteFuse, 
                  gg.CD4_Tcells.ADT, 
                  gg.CD4_Tcells.RNA, ncol=3) %>% ggsave(filename="Figure2d.pdf", width=15, height=5.5)
```

# Plot UMAP of CD4+ T-cells highlighted by different marker genes

```{r}
memory.markers <- c("hg19_S100A4", "hg19_CRIP1", "hg19_AHNAK")
naive.markers <- c("hg19_TCF", "hg19_ID3", "hg19_CCR7", "hg19_SELL")
DR.markers <- c("hg19_GNLY", "hg19_GZMB",  "hg19_PRF1", "hg19_GZMA", "hg19_NKG7", "hg19_HLA-DRB1", "hg19_HLA-DPA1")
```

```{r plot_markers, fig.height=5, fig.width=20, fig.cap="UMAP of CD4+ T-cells highlighted by marker gene expression."}
gg.CD4_Tcells.CD4 <- 
    ggplot(dftoPlot, aes(x = UMAP1, y = UMAP2, colour = as.matrix(clr_adt_mat_control.filtered)["CD4",])) +
    geom_point(alpha = 0.5) +
    coord_cartesian(xlim=c(8,15), ylim=c(-7,3)) + 
    scale_color_viridis_c() +
    theme_bw() +
    ggtitle("CD4 receptor expression") + 
    theme(aspect.ratio = 1, legend.position = "none") 

gg.CD4_Tcells.CD27 <- 
    ggplot(dftoPlot, aes(x = UMAP1, y = UMAP2, colour = as.matrix(clr_adt_mat_control.filtered)["CD27",])) +
    geom_point(alpha = 0.5) +
    coord_cartesian(xlim=c(8,15), ylim=c(-7,3)) + 
    scale_color_viridis_c() +
    theme_bw() +
    ggtitle("CD27 receptor expression") + 
    theme(aspect.ratio = 1, legend.position = "none") 

gg.CD4_Tcells.memory <- 
    ggplot(dftoPlot, aes(x = UMAP1, y = UMAP2, 
                         colour = colMeans(normalize.quantiles(logcounts(sce_control))[rownames(sce_control) %in% c(memory.markers),]))) +
    geom_point(alpha = 0.5) +
    coord_cartesian(xlim=c(8,15), ylim=c(-7,3)) + 
    scale_color_viridis_c() +
    theme_bw() +
    ggtitle("Memory cell marker score") + 
    theme(aspect.ratio = 1, legend.position = "none") 

gg.CD4_Tcells.naive <- 
    ggplot(dftoPlot, aes(x = UMAP1, y = UMAP2, 
                         colour = colMeans(normalize.quantiles(logcounts(sce_control))[rownames(sce_control) %in% c(naive.markers),]))) +
    geom_point(alpha = 0.5) +
    coord_cartesian(xlim=c(8,15), ylim=c(-7,3)) + 
    scale_color_viridis_c() +
    theme_bw() +
    ggtitle("Naive cell marker score") + 
    theme(aspect.ratio = 1, legend.position = "none") 

gg.CD4_Tcells.DR <- 
    ggplot(dftoPlot, aes(x = UMAP1, y = UMAP2, 
                         colour = colMeans(normalize.quantiles(logcounts(sce_control))[rownames(sce_control) %in% c(DR.markers),]))) +
    geom_point(alpha = 0.5) +
    coord_cartesian(xlim=c(8,15), ylim=c(-7,3)) + 
    scale_color_viridis_c() +
    theme_bw() +
    ggtitle("CD27- DR+ T-cell marker score") + 
    theme(aspect.ratio = 1, legend.position = "none") 

ggpubr::ggarrange(gg.CD4_Tcells.CD4, 
                  gg.CD4_Tcells.CD27,
                  gg.CD4_Tcells.memory,
                  gg.CD4_Tcells.naive,
                  gg.CD4_Tcells.DR, 
                  ncol=5) 

ggpubr::ggarrange(gg.CD4_Tcells.CD4, 
                  gg.CD4_Tcells.CD27,
                  gg.CD4_Tcells.memory,
                  gg.CD4_Tcells.naive,
                  gg.CD4_Tcells.DR, 
                  ncol=5)  %>% ggsave(filename="Figure2e.pdf", width=25, height=6)
```

