---
title: "ligand_receptor_scatterplot"
author: "Hani Jieun Kim"
date: "11/03/2020"
output: 
  BiocStyle::html_document:
    code_folding: hide
    number_sections: yes
    toc: yes  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#  Ligand and receptor scatterplot

The script reproduces the scatter plot in Figure 5d. It is shown in Figure \@ref(fig:plot).
```{r load, echo=FALSE}
setwd("/Users/hani/Dropbox (Sydney Uni)/Single Cell Reserach/SCProj-CITEseq/codes/Code/CodeRepository")

library(ggplot2)
library(ggthemes)
library(reshape2)
library(ggpubr)
library(tidyverse)
```

Load expression matrices
```{r}
#read in log-transformed RNA expression matrix
rna_log_norm <- readRDS("lr_interaction_rna_log_norm.rds")
head(rna_log_norm[1:10,1:10])
#read in log-transformed and normalised ADT count matrix
adt_log_norm <- readRDS("lr_interaction_adt_log_norm.rds")
head(adt_log_norm[1:10,1:10])
```

# Ligand-receptor predictions

Load ligand-receptor pairs identified as significantly interacting by CiteFuse
```{r}
lr_pairs_CiteFuse <- readRDS("lr_pairs_CiteFuse.rds") # used `se_alt_exp = TRUE` to run `ligandReceptorTest`
lr_pairs_RNA <- readRDS("lr_pairs_RNA.rds") # used `se_alt_exp = FALSE` to run `ligandReceptorTest`
```

Note that for each pair of interactions (cluster A_cluster B), cluster A is enriched by the ligand and cluster B is enriched by the receptor. 
```{r}
head(lr_pairs_CiteFuse)
```

# Categories of ligand-receptor interactions

Function to substitute gene names with protein names and vice versa
```{r}
gsub_rna <- function(x) {
  x <- gsub("SELL", "CD62L", x)
  x <- gsub("ITGAM", "CD11b", x)
  x <- gsub("FCGR3A", "CD16", x)
  x <- gsub("FCGR3B", "CD16", x)
  x <- gsub("PTPRC", "CD45", x)
  x <- gsub("DPP4", "CD26 (Adenosine)", x)
  return(x)
}

gsub_rna_backward <- function(x) {
  x <- gsub("CD62L", "SELL", x)
  x <- gsub("CD11b", "ITGAM", x)
  x <- gsub("CD16", "FCGR3B", x)
  x2 <- gsub("FCGR3B", "FCGR3A", x)
  x <- unique(x, x2)
  x <- gsub("CD45", "PTPRC", x)
  x <- gsub("CD26 \\(Adenosine)", "DPP4", x)
  return(x)
}
```

Generate the three categories of ligand-receptor interactions (common, CiteFuse-specific, and RNA-specific)
```{r}
lr_pairs_RNA_alias <- lapply(lr_pairs_RNA, gsub_rna)

interactions <- lapply(1:length(lr_pairs_RNA_alias), function(x) {
  rna_tmp <- lr_pairs_RNA_alias[[x]]
  adt_tmp <- lr_pairs_CiteFuse[[x]]
  common <- intersect(rna_tmp, adt_tmp)
  adt_specific <- adt_tmp[!adt_tmp %in% common]
  rna_specific <- rna_tmp[!rna_tmp %in% common]
  
  c(RNA_specific = length(rna_specific), CiteFuse_specific = length(adt_specific), Common = length(common))
})

interactions <- do.call(rbind, interactions)
rownames(interactions) <- names(lr_pairs_CiteFuse)
head(interactions)
```


# Identify receptors by cluster

Identify the receptors involved in a ligand-receptor interaction by cluster
```{r}
lr_pairs_Receptors <- lapply(1:length(lr_pairs_RNA_alias), function(x) {
  
  rna_tmp <- lr_pairs_RNA_alias[[x]]
  adt_tmp <- lr_pairs_CiteFuse[[x]]
  common <- intersect(rna_tmp, adt_tmp)
  adt_specific <- adt_tmp[!adt_tmp %in% common]
  rna_specific <- rna_tmp[!rna_tmp %in% common]
  
  res <- list(common = common, ADT = adt_specific, RNA = rna_specific)
  res <- lapply(res, function(x) {
    unique(sapply(strsplit(x, "_"), "[[", 2))
  })
  return(res)
})

#### extract receptors
names(lr_pairs_Receptors) <- sapply(strsplit(names(lr_pairs_RNA_alias), "_"), "[[", 2) 

#### group receptors by cluster
lr_pairs_Receptors_byCluster <- lapply(1:16, function(i) {
  tmp <- lr_pairs_Receptors[names(lr_pairs_Receptors) == i]
  
  common <- unique(unlist(lapply(tmp, "[[", "common")))
  ADT <- unique(unlist(lapply(tmp, "[[", "ADT")))
  RNA <- unique(unlist(lapply(tmp, "[[", "RNA")))
  
  res <- list(Common = common, CiteFuse_specific = ADT, RNA_specific = RNA)
  return(res)
})
```

# Calculate ranked expression of receptors by cluster

Load cluster labels from CiteFuse
```{r}
load("cluster_labels.RData", verbose=T)
```

First calculate average expression of RNA and protein of receptors by cluster 
```{r}
receptor_meanExprs_RNA <- apply(rna_log_norm, 1, function(x) {
  aggregate(x, list(CiteFuse_cluster_labels), mean)$x
})

receptor_meanExprs_ADT <- apply(adt_log_norm, 1, function(x) {
  aggregate(x, list(CiteFuse_cluster_labels), mean)$x
})
```

Second rank the average expression across the clusters
```{r}
receptor_rankedExprs_RNA  <- apply(receptor_meanExprs_RNA, 2, function(x){
  rank(x, ties.method="random")
})

receptor_rankedExprs_ADT <- apply(receptor_meanExprs_ADT, 2, function(x){
  rank(x, ties.method="random")
})

receptor_rankedExprs_RNA <- t(receptor_rankedExprs_RNA)
receptor_rankedExprs_ADT <- t(receptor_rankedExprs_ADT)

receptor_ranking_RNA <- receptor_ranking_ADT <- list()

for (i in 1:length(lr_pairs_Receptors_byCluster)) {
  
  Common <- lr_pairs_Receptors_byCluster[[i]]$Common
  CiteFuse_specific <- lr_pairs_Receptors_byCluster[[i]]$CiteFuse_specific
  RNA_specific <- lr_pairs_Receptors_byCluster[[i]]$RNA_specific
  
  Common.RNA <- receptor_rankedExprs_RNA[gsub_rna_backward(Common), i]
  CiteFuse_specific.RNA <- receptor_rankedExprs_RNA[gsub_rna_backward(CiteFuse_specific), i]
  RNA_specific.RNA <- receptor_rankedExprs_RNA[gsub_rna_backward(RNA_specific), i]
  
  Common.ADT <- receptor_rankedExprs_ADT[Common, i]
  CiteFuse_specific.ADT <- receptor_rankedExprs_ADT[CiteFuse_specific, i]
  RNA_specific.ADT <- receptor_rankedExprs_ADT[RNA_specific, i]
  
  
  receptor_ranking_RNA[[i]] <- list(RNA_specific = RNA_specific.RNA,
                                    CiteFuse_specific = CiteFuse_specific.RNA,
                                    Common = Common.RNA)
  
  receptor_ranking_ADT[[i]] <- list(RNA_specific = RNA_specific.ADT,
                                    CiteFuse_specific = CiteFuse_specific.ADT,
                                    Common = Common.ADT)
}

RNA_specific.RNA <- unlist(lapply(receptor_ranking_RNA, "[[", "RNA_specific"))
CiteFuse_specific.RNA <- unlist(lapply(receptor_ranking_RNA, "[[", "CiteFuse_specific"))
Common.RNA <- unlist(lapply(receptor_ranking_RNA, "[[", "Common"))

RNA_specific.ADT <- unlist(lapply(receptor_ranking_ADT, "[[", "RNA_specific"))
CiteFuse_specific.ADT <- unlist(lapply(receptor_ranking_ADT, "[[", "CiteFuse_specific"))
Common.ADT <- unlist(lapply(receptor_ranking_ADT, "[[", "Common"))
```

# Plot of relative ranking of RNA versus protein expression of the receptors
Geneate a scatter of RNA and protein expression of receptors ranked across the clusters. Scatter plot is generated for each category of receptors in Figure \@ref(fig:plot))
```{r}

pdf(file="Figure5d.pdf", width=12, height=5)
par(mfrow=c(1,3))
plot(RNA_specific.ADT, RNA_specific.RNA, 
     col="#B3DE69", 
     pch=19, 
     xlim=c(1,16), ylim=c(1,16),
     xlab="Relative ranking of protein \nexpression across clusters",
     ylab="Relative ranking of RNA expression across clusters",
     main="RNA-specific")
abline(lm(RNA_specific.RNA~RNA_specific.ADT), col="#B3DE69", lty="dotted", lwd=2)
plot(CiteFuse_specific.ADT, CiteFuse_specific.RNA, 
     col="#FDB462", 
     pch=19, 
     xlim=c(1,16), ylim=c(1,16),
     xlab="Relative ranking of protein \nexpression across clusters",
     ylab="Relative ranking of RNA expression across clusters",
     main="CiteFuse_specific")
abline(lm(CiteFuse_specific.RNA~CiteFuse_specific.ADT), col="#FDB462", lty="dotted", lwd=2)
plot(Common.ADT, Common.RNA, 
     col="#FCCDE5", 
     pch=19, 
     xlab="Relative ranking of protein \nexpression across clusters",
     ylab="Relative ranking of RNA expression across clusters",
     xlim=c(1,16), ylim=c(1,16),
     main="Common")
abline(lm(Common.RNA~Common.ADT), col="#FCCDE5", lty="dotted", lwd=2)
dev.off()
```

```{r plot, fig.cap="Relative ranking of RNA vs protein expression of receptors across clusters", fig.height=5, fig.width=10}

par(mfrow=c(1,3))
plot(RNA_specific.ADT, RNA_specific.RNA, 
     col="#B3DE69", 
     pch=19, 
     xlim=c(1,16), ylim=c(1,16),
     xlab="Relative ranking of protein \nexpression across clusters",
     ylab="Relative ranking of RNA expression across clusters",
     main="RNA-specific")
abline(lm(RNA_specific.RNA~RNA_specific.ADT), col="#B3DE69", lty="dotted", lwd=2)
plot(CiteFuse_specific.ADT, CiteFuse_specific.RNA, 
     col="#FDB462", 
     pch=19, 
     xlim=c(1,16), ylim=c(1,16),
     xlab="Relative ranking of protein \nexpression across clusters",
     ylab="Relative ranking of RNA expression across clusters",
     main="CiteFuse_specific")
abline(lm(CiteFuse_specific.RNA~CiteFuse_specific.ADT), col="#FDB462", lty="dotted", lwd=2)
plot(Common.ADT, Common.RNA, 
     col="#FCCDE5", 
     pch=19, 
     xlab="Relative ranking of protein \nexpression across clusters",
     ylab="Relative ranking of RNA expression across clusters",
     xlim=c(1,16), ylim=c(1,16),
     main="Common")
abline(lm(Common.RNA~Common.ADT), col="#FCCDE5", lty="dotted", lwd=2)

```

Assess correlation between RNA and protein expression of receptors by cateory
```{r}
cor.test(RNA_specific.ADT, RNA_specific.RNA)
cor.test(CiteFuse_specific.ADT, CiteFuse_specific.RNA)
cor.test(Common.ADT, Common.RNA)
```


#Session Info
```{r sessinf}
sessionInfo()
```
