---
title: "ligand_receptor_expression"
author: "Hani Jieun Kim"
date: "11/03/2020"
output: 
  BiocStyle::html_document:
    code_folding: hide
    number_sections: yes
    toc: yes  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#  Ligand and receptor expression

The script reproduces the barplot in Figure 5c. It is shown in Figure \@ref(fig:plot).
```{r load, echo=FALSE}
setwd("/Users/hani/Dropbox (Sydney Uni)/Single Cell Reserach/SCProj-CITEseq/codes/Code/CodeRepository")

library(ggplot2)
library(ggthemes)
library(reshape2)
library(ggpubr)
library(tidyverse)
```

Load expression matrices
```{r}
#read in log-transformed RNA expression matrix
rna_log_norm <- readRDS("lr_interaction_rna_log_norm.rds")
head(rna_log_norm[1:10,1:10])
#read in log-transformed and normalised ADT count matrix
adt_log_norm <- readRDS("lr_interaction_adt_log_norm.rds")
head(adt_log_norm[1:10,1:10])
```

# Ligand-receptor predictions

Load ligand-receptor pairs identified as significantly interacting by CiteFuse
```{r}
lr_pairs_CiteFuse <- readRDS("lr_pairs_CiteFuse.rds") # used `se_alt_exp = TRUE` to run `ligandReceptorTest`
lr_pairs_RNA <- readRDS("lr_pairs_RNA.rds") # used `se_alt_exp = FALSE` to run `ligandReceptorTest`
```

Note that for each pair of interactions (cluster A_cluster B), cluster A is enriched by the ligand and cluster B is enriched by the receptor. 
```{r}
head(lr_pairs_CiteFuse)
```

# Categories of ligand-receptor interactions

Function to substitute gene names with protein names and vice versa
```{r}
gsub_rna <- function(x) {
  x <- gsub("SELL", "CD62L", x)
  x <- gsub("ITGAM", "CD11b", x)
  x <- gsub("FCGR3A", "CD16", x)
  x <- gsub("FCGR3B", "CD16", x)
  x <- gsub("PTPRC", "CD45", x)
  x <- gsub("DPP4", "CD26 (Adenosine)", x)
  return(x)
}

gsub_rna_backward <- function(x) {
  x <- gsub("CD62L", "SELL", x)
  x <- gsub("CD11b", "ITGAM", x)
  x <- gsub("CD16", "FCGR3B", x)
  x2 <- gsub("FCGR3B", "FCGR3A", x)
  x <- unique(x, x2)
  x <- gsub("CD45", "PTPRC", x)
  x <- gsub("CD26 \\(Adenosine)", "DPP4", x)
  #x <- gsub( "PD-1 \\(CD279)", "PDCD1", x)
  return(x)
}
```

Generate the three categories of ligand-receptor interactions (common, CiteFuse-specific, and RNA-specific)
```{r}
lr_pairs_RNA_alias <- lapply(lr_pairs_RNA, gsub_rna)

interactions <- lapply(1:length(lr_pairs_RNA_alias), function(x) {
  rna_tmp <- lr_pairs_RNA_alias[[x]]
  adt_tmp <- lr_pairs_CiteFuse[[x]]
  common <- intersect(rna_tmp, adt_tmp)
  adt_specific <- adt_tmp[!adt_tmp %in% common]
  rna_specific <- rna_tmp[!rna_tmp %in% common]
  
  c(RNA_specific = length(rna_specific), CiteFuse_specific = length(adt_specific), Common = length(common))
})

interactions <- do.call(rbind, interactions)
rownames(interactions) <- names(lr_pairs_CiteFuse)
head(interactions)
```


#  Identify receptors by cluster

Identify the receptors involved in a ligand-receptor interaction by cluster
```{r}
lr_pairs_Receptors <- lapply(1:length(lr_pairs_RNA_alias), function(x) {
  
  rna_tmp <- lr_pairs_RNA_alias[[x]]
  adt_tmp <- lr_pairs_CiteFuse[[x]]
  common <- intersect(rna_tmp, adt_tmp)
  adt_specific <- adt_tmp[!adt_tmp %in% common]
  rna_specific <- rna_tmp[!rna_tmp %in% common]
  
  res <- list(common = common, ADT = adt_specific, RNA = rna_specific)
  res <- lapply(res, function(x) {
    unique(sapply(strsplit(x, "_"), "[[", 2))
  })
  return(res)
})

#### extract receptors
names(lr_pairs_Receptors) <- sapply(strsplit(names(lr_pairs_RNA_alias), "_"), "[[", 2) 

#### group receptors by cluster
lr_pairs_Receptors_byCluster <- lapply(1:16, function(i) {
  tmp <- lr_pairs_Receptors[names(lr_pairs_Receptors) == i]
  
  common <- unique(unlist(lapply(tmp, "[[", "common")))
  ADT <- unique(unlist(lapply(tmp, "[[", "ADT")))
  RNA <- unique(unlist(lapply(tmp, "[[", "RNA")))
  
  res <- list(Common = common, CiteFuse_specific = ADT, RNA_specific = RNA)
  return(res)
})
```

#  Boxplot to compare RNA and protein expression of receptors

Load cluster labels from CiteFuse
```{r}
load("cluster_labels.RData", verbose=T)
```

Extract RNA and protein expression of receptors for each cluster
```{r}
receptor_expression_RNA <- receptor_expression_ADT <- list()

for(i in 1:length(lr_pairs_Receptors_byCluster)) {
  
  Common <- lr_pairs_Receptors_byCluster[[i]]$Common
  CiteFuse_specific <- lr_pairs_Receptors_byCluster[[i]]$CiteFuse_specific
  RNA_specific <- lr_pairs_Receptors_byCluster[[i]]$RNA_specific
  
  
  Common.RNA <- rna_log_norm[gsub_rna_backward(Common), CiteFuse_cluster_labels == i]
  CiteFuse_specific.RNA <- rna_log_norm[gsub_rna_backward(CiteFuse_specific), CiteFuse_cluster_labels == i]
  RNA_specific.RNA <- rna_log_norm[gsub_rna_backward(RNA_specific), CiteFuse_cluster_labels == i]
  
  
  Common.ADT <- adt_log_norm[(Common), CiteFuse_cluster_labels == i]
  CiteFuse_specific.ADT <- adt_log_norm[(CiteFuse_specific), CiteFuse_cluster_labels == i]
  RNA_specific.ADT <- adt_log_norm[(RNA_specific), CiteFuse_cluster_labels == i]
  
  
   res <- list(RNA_specific = c(RNA_specific.ADT),
              CiteFuse_specific = c(CiteFuse_specific.ADT),
              Common = c(Common.ADT))
  
  receptor_expression_ADT[[i]] <- melt(res)
  receptor_expression_ADT[[i]][, 2] <- factor(as.character(receptor_expression_ADT[[i]][, 2]), levels = names(res))
  colnames(receptor_expression_ADT[[i]]) <- c("value", "method")
  
  res <- list(RNA_specific = c(RNA_specific.RNA),
              CiteFuse_specific = c(CiteFuse_specific.RNA),
              Common = c(Common.RNA))
  
  receptor_expression_RNA[[i]] <- melt(res)
  receptor_expression_RNA[[i]][, 2] <- factor(as.character(receptor_expression_RNA[[i]][, 2]), levels = names(res))
  colnames(receptor_expression_RNA[[i]]) <- c("value", "method")
  
}


exprs_ADT <- do.call(rbind, receptor_expression_ADT)
exprs_RNA <- do.call(rbind, receptor_expression_RNA)
```

Geneate boxplot of RNA and ADT expression of receptors identified in a ligand-receptor interaction by category (Figure \@ref(fig:plot))
```{r plot, fig.cap="Expression of RNA and ADT of receptors by category."}
gg_ADT <- ggplot(exprs_ADT, aes(x = method, y = value, fill = method)) +
    geom_boxplot(notch = F, notchwidth = 0.3, alpha = 0.8) + 
    scale_x_discrete("method", breaks = factor(levels(exprs_ADT$method)), drop=FALSE) +
    ylab("Normalised log expression") +
    theme_classic() +
    ggtitle("Expression of ADT") +
    scale_fill_manual(values=(c("#B3DE69","#FDB462", "#FCCDE5"))) + 
    theme(axis.text.x = element_blank())
gg_RNA <- ggplot(exprs_RNA, aes(x = method, y = value, fill = method)) +
    geom_boxplot(notch = F, notchwidth = 0.3, alpha = 0.8) + 
    scale_x_discrete("method", breaks = factor(levels(exprs_RNA$method)), drop=FALSE) +
    ylab("Normalised log expression") +
    theme_classic() +
    ggtitle("Expression of RNA") +
    scale_fill_manual(values=(c("#B3DE69","#FDB462", "#FCCDE5"))) + 
    theme(axis.text.x = element_blank())

ggarrange(gg_RNA, gg_ADT, ncol=2, common.legend = T) 

ggarrange(gg_RNA, gg_ADT, ncol=2, common.legend = T) %>% ggsave(file="Figure5c.pdf", width = 5, height = 5)
```

#Session Info
```{r sessinf}
sessionInfo()
```