---
title: "ligand_receptor_interactions"
author: "Hani Jieun Kim"
date: "11/03/2020"
output: 
  BiocStyle::html_document:
    code_folding: hide
    number_sections: yes
    toc: yes  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Ligand-receptor interactions

The script reproduces the barplot in Figure 5a. It is shown in Figure \@ref(fig:plot)
```{r load, echo=FALSE}
setwd("/Users/hani/Dropbox (Sydney Uni)/Single Cell Reserach/SCProj-CITEseq/codes/Code/CodeRepository")

library(ggplot2)
library(ggthemes)
library(reshape2)
library(tidyverse)
```

Load expression matrices
```{r}
#read in log-transformed RNA expression matrix
rna_log_norm <- readRDS("lr_interaction_rna_log_norm.rds")
head(rna_log_norm[1:10,1:10])
#read in log-transformed and normalised ADT count matrix
adt_log_norm <- readRDS("lr_interaction_adt_log_norm.rds")
head(adt_log_norm[1:10,1:10])
```

# Ligand-receptor predictions
Load ligand-receptor pairs identified as significantly interacting by CiteFuse
```{r}
lr_pairs_CiteFuse <- readRDS("lr_pairs_CiteFuse.rds") # used `se_alt_exp = TRUE` to run `ligandReceptorTest`
lr_pairs_RNA <- readRDS("lr_pairs_RNA.rds") # used `se_alt_exp = FALSE` to run `ligandReceptorTest`
```

Note that for each pair of interactions (cluster A_cluster B), cluster A is enriched by the ligand and cluster B is enriched by the receptor. 
```{r}
head(lr_pairs_CiteFuse)
```

# Categories of ligand-receptor interactions

Function to substitute gene names with protein names
```{r}
gsub_rna <- function(x) {
  x <- gsub("SELL", "CD62L", x)
  x <- gsub("ITGAM", "CD11b", x)
  x <- gsub("FCGR3A", "CD16", x)
  x <- gsub("FCGR3B", "CD16", x)
  x <- gsub("PTPRC", "CD45", x)
  x <- gsub("DPP4", "CD26 (Adenosine)", x)
  return(x)
}
```

Generate the three categories of ligand-receptor interactions (common, CiteFuse-specific, and RNA-specific)
```{r}
lr_pairs_RNA_alias <- lapply(lr_pairs_RNA, gsub_rna)

interactions <- lapply(1:length(lr_pairs_RNA_alias), function(x) {
  rna_tmp <- lr_pairs_RNA_alias[[x]]
  adt_tmp <- lr_pairs_CiteFuse[[x]]
  common <- intersect(rna_tmp, adt_tmp)
  adt_specific <- adt_tmp[!adt_tmp %in% common]
  rna_specific <- rna_tmp[!rna_tmp %in% common]
  
  c(RNA_specific = length(rna_specific), CiteFuse_specific = length(adt_specific), Common = length(common))
})

interactions <- do.call(rbind, interactions)
rownames(interactions) <- names(lr_pairs_CiteFuse)
head(interactions)
```

# Barplot of total number of interactions by category
The number of ligand-receptor interactions indentified in each category are shown in Figure \@ref(fig:plot)
```{r plot, fig.cap="Number of interactions identified by each category."}
df <- melt(interactions)
colnames(df) <- c("LR_cluster", "Class", "Value")
df$Receptor_cluster <- unlist(lapply(strsplit(as.character(df$LR_cluster),"_"), "[[", 2))
df$Receptor_cluster <- factor(df$Receptor_cluster, levels = unique(df$Receptor_cluster))

interactions.barplot <- ggplot(df, aes(x = Receptor_cluster, y = Value, fill = Class)) +
    geom_bar(stat = "identity") +
    ylab("Number of ligand-receptor interactions") +
    xlab("Clusters") +
    theme_bw() +
    scale_fill_manual(values=c("#B3DE69","#FDB462", "#FCCDE5")) + 
    theme(axis.text.x = element_text(angle = 90, size = 8))

interactions.barplot 

ggsave(interactions.barplot, file="Figure5b.pdf", width=10, height=10)
```

#Session Info
```{r sessinf}
sessionInfo()
```