---
title: "PCA_doublet_filtering"
author: "Hani Jieun Kim"
date: "11/03/2020"
output: 
  BiocStyle::html_document:
    code_folding: hide
    number_sections: yes
    toc: yes  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# PCA plot of HTO count 

The script reproduces the PCA plot of HTO count in Figure 3c. It is shown in Figure \@ref(fig:plot).
```{r load, echo=FALSE}
setwd("/Users/hani/Dropbox (Sydney Uni)/Single Cell Reserach/SCProj-CITEseq/codes/Code/CodeRepository")

library(ggplot2)
library(ggthemes)
library(reshape2)
library(ggpubr)
library(tidyverse)
```

Load the unfiltered HTO count matrix and perform log transformation
```{r}
load("Unfiltered.RData", verbose=T)
HTO_logcounts <- log(hto_control+1)
```

# Filter doublets from the HTO count matrix

Use doublet labels identified from CiteFuse, Scrublet, HTODemux, and DoubletFinder to filter the HTO count matrix.
```{r}
load("doublet_labels.Rdata", verbose = T)
load("cellHashing_GT.RData", verbose = T)

runPCA <- function(labels) {
    
    HTO_pca <- prcomp(as.matrix(t(HTO_logcounts[,labels==FALSE])), scale = T)
    
    HTO_pca_toPlot <- data.frame(
        PCA1 = HTO_pca$x[, 1],
        PCA2 = HTO_pca$x[, 2]
    )
    
    return(HTO_pca_toPlot)
}

HTO_pca_toPlot <- lapply(doublet.labels, runPCA)
```

# Plot HTO count filtered by each doublet detection method
```{r plot, fig.cap="PCA plot of HTO counts filtered for doublets.", fig.height=5, fig.width=20}
gg.HTO_unfiltered <- ggplot(HTO_pca_toPlot[[1]], aes(x = PCA1, y = PCA2, col = cellHashinglabel[doublet.labels$Unfiltered == F])) +
    geom_point(alpha=0.5) +
    theme_classic() +
    scale_color_manual(values=c("#80B1D3","#FDB462","#FB8072", "#8DD3C7","#B3DE69")) + 
    theme(aspect.ratio = 1,  legend.position = "bottom") +
    labs(title = "Unfiltered")
gg.HTO_CiteFuse <- ggplot(HTO_pca_toPlot[[2]], aes(x = PCA1, y = PCA2, col = cellHashinglabel[doublet.labels$CiteFuse == F])) +
    geom_point(alpha=0.5) +
    theme_classic() +
    scale_color_manual(values=c("#80B1D3","#FDB462","#FB8072", "#8DD3C7","#B3DE69")) + 
    theme(aspect.ratio = 1,  legend.position = "bottom") +
    labs(title = "CiteFuse")
gg.HTO_Scrublet <- ggplot(HTO_pca_toPlot[[3]], aes(x = PCA1, y = PCA2, col = cellHashinglabel[doublet.labels$Scrublet == F])) +
    geom_point(alpha=0.5) +
    theme_classic() +
    scale_color_manual(values=c("#80B1D3","#FDB462","#FB8072", "#8DD3C7","#B3DE69")) + 
    theme(aspect.ratio = 1,  legend.position = "bottom") +
    labs(title = "Scrublet")
gg.HTO_HTODemux <- ggplot(HTO_pca_toPlot[[4]], aes(x = PCA1, y = PCA2, col = cellHashinglabel[doublet.labels$HTODemux == F])) +
    geom_point(alpha=0.5) +
    theme_classic() +
    scale_color_manual(values=c("#80B1D3","#FDB462","#FB8072", "#8DD3C7","#B3DE69")) + 
    theme(aspect.ratio = 1,  legend.position = "bottom") +
    labs(title = "HTODemux")
gg.HTO_DoubletFinder <- ggplot(HTO_pca_toPlot[[5]], aes(x = PCA1, y = PCA2, col = cellHashinglabel[doublet.labels$DoubletFinder == F])) +
    geom_point(alpha=0.5) +
    theme_classic() +
    scale_color_manual(values=c("#80B1D3","#FDB462","#FB8072", "#8DD3C7","#B3DE69")) + 
    theme(aspect.ratio = 1,  legend.position = "bottom") +
    labs(title = "DoubletFinder")

ggpubr::ggarrange(gg.HTO_unfiltered,
                  gg.HTO_HTODemux,
                  gg.HTO_Scrublet,
                  gg.HTO_DoubletFinder,
                  gg.HTO_CiteFuse,
                  ncol=5,
                  common.legend = T)


ggpubr::ggarrange(gg.HTO_unfiltered,
                  gg.HTO_HTODemux,
                  gg.HTO_Scrublet,
                  gg.HTO_DoubletFinder,
                  gg.HTO_CiteFuse,
                  ncol=5,
                  common.legend = T) %>% ggsave(filename="Figure3c.pdf", width = 20, height = 5)
```

#Session Info
```{r sessinf}
sessionInfo()
```