---
title: "purity_plot"
author: "Hani Jieun Kim"
date: "11/03/2020"
output: 
  BiocStyle::html_document:
    code_folding: hide
    number_sections: yes
    toc: yes  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#  Purity plot 

The script reproduces the purity plot in Figure 4b. It is shown in Figure \@ref(fig:plot).
```{r load, echo=FALSE}

setwd("/Users/hani/Dropbox (Sydney Uni)/Single Cell Reserach/SCProj-CITEseq/codes/Code/CodeRepository")

library(ggplot2)
library(ggthemes)
library(reshape2)
library(ggpubr)
library(tidyverse)
library(mixtools)
source("functions.R")

```

# Mixture modelling for CD4 and CD8 populations

Fit a mixture model on CD4/CD8 and CD11c expression and calculate threshold. 
```{r}
load("doublet_labels.RData", verbose=T)
load("Unfiltered.RData", verbose=T)



CD4_pass_labels <- CD8_pass_labels <- list()



for (i in 1:length(doublet.labels)) {
  
  adt_count <- adt_control[,doublet.labels[[i]] == FALSE]
  adt_logcount <- log(adt_count + 1)
  logcount_CD4 <- adt_logcount[rownames(adt_logcount) %in% c("CD4","CD11c"),]
  logcount_CD8 <- adt_logcount[rownames(adt_logcount) %in% c("CD8","CD11c"),]
  
  mixmdl_CD4 <- list()
  par(mfrow = c(2,2))
  for (gene in 1:nrow(logcount_CD4)) {
    mixmdl_CD4[[gene]] <- try(mixtools::normalmixEM(logcount_CD4[gene,][logcount_CD4[gene,]>0], 
                                                    fast = T, maxrestarts = 1000,
                                                    k = 2, maxit = 10000,
                                                    mu = c(0, 8),
                                                    ECM = TRUE, verb = F),
                              silent = T)
    plot(mixmdl_CD4[[gene]], whichplots = 2)
  }
  
  CD4_threshold <- lapply(mixmdl_CD4, getThreshold)
  names(CD4_threshold) <- NULL
  CD4_threshold <- unlist(CD4_threshold)
  
  print(paste0("CD4 threshold for ", names(doublet.labels)[[i]], "-filtered matrix"))
  print(CD4_threshold)
  
  CD4_pass <- sapply(1:nrow(logcount_CD4), function(x) {
    logcount_CD4[x,] > CD4_threshold[x]
  })
  table(rowSums(CD4_pass))
  colnames(CD4_pass) <- rownames(logcount_CD4)
  
  CD4_pass_labels[[i]] <- 
    
    apply(CD4_pass, 1, function(x) {
      if (sum(x) == 0) {
        "negative"
      } else if (sum(x) == 1) {
        "CD4"
      } else {
        "negative"
      }
    })
  
  mixmdl_CD8 <- list()
  par(mfrow = c(2,2))
  for (gene in 1:nrow(logcount_CD8)) {
    mixmdl_CD8[[gene]] <- try(mixtools::normalmixEM(logcount_CD8[gene,][logcount_CD8[gene,]>0], 
                                                    fast = T, maxrestarts = 1000,
                                                    k = 2, maxit = 10000,
                                                    mu = c(0, 8),
                                                    ECM = TRUE, verb = F),
                              silent = T)
    plot(mixmdl_CD8[[gene]], whichplots = 2)
  }
  
  CD8_threshold <- lapply(mixmdl_CD8, getThreshold)
  names(CD8_threshold) <- NULL
  CD8_threshold <- unlist(CD8_threshold)
  
  print(paste0("CD8 threshold for ", names(doublet.labels)[[i]], "-filtered matrix"))
  print(CD8_threshold)
  
  CD8_pass <- sapply(1:nrow(logcount_CD8), function(x) {
    logcount_CD8[x,] > CD8_threshold[x]
  })
  table(rowSums(CD8_pass))
  colnames(CD8_pass) <- rownames(logcount_CD8)
  
  CD8_pass_labels[[i]] <- 
    
    sapply(CD8_pass[,2], function(x) {
      if (sum(x) == 0) {
        "negative"
      } else if (sum(x) == 1) {
        "CD8"
      } else {
        "negative"
      }
    })
  
}
```

Use mixture model to determine CD4 and CD8 positive cells for the five dataset, each filtered using a different doublet detection method.
```{r}
lapply(CD4_pass_labels, table)
lapply(CD8_pass_labels, table)
```

# Calculate purity score

Calculate the purity of a cluster for either CD4 or CD8 T-cells. 
```{r}
load("cluster_labels_doubletfiltered.RData", verbose = T)
load("annotation_colors.RData")

CD4clusterPurity <- function(clusters, labels) {
  sapply(1:16, function(x) {
    
    max(table(clusters,labels)[x,"CD4"]) / sum(table(clusters,labels)[x,]) 
    
    })
  
}

CD8clusterPurity <- function(clusters, labels) {
  sapply(1:16, function(x) {
    max(table(clusters,labels)[x,"CD8"]) / sum(table(clusters,labels)[x,]) })
}


CD4purityScore <- lapply(1:5, function(x) {
  CD4clusterPurity(cluster_labels[[x]], CD4_pass_labels[[x]])
})

CD8purityScore <- lapply(1:5, function(x) {
  CD8clusterPurity(cluster_labels[[x]], CD8_pass_labels[[x]])
})



CD4_dat <- do.call(cbind, CD4purityScore)
rownames(CD4_dat) <- c(1:16)
CD4_dat<- reshape2::melt(CD4_dat)
colnames(CD4_dat) <- c("cluster", "method", "purity")

CD8_dat <- do.call(cbind, CD8purityScore)
rownames(CD8_dat) <- c(1:16)
CD8_dat<- reshape2::melt(CD8_dat)
colnames(CD8_dat) <- c("cluster", "method", "purity")

combined <- cbind(CD4_dat, CD8_dat[,3])
colnames(combined) <- c("cluster", "method", "purity_CD4", "purity_CD8")

combined$method <- as.factor(combined$method)
combined$cluster <- as.factor(combined$cluster)
```



```{r plot, fig.cap="Purity scores of CD8+ T cells against CD4+ T-cells", fig.height=5, fig.width=20}

purity.unfiltered <- ggplot(combined[combined$method == 1,], 
                            aes(x=purity_CD4, y=purity_CD8, col=cluster)) +
  geom_point(size=4) +
  geom_hline(yintercept = 0.1, col="red", lty="dotted", lwd=1) + 
  geom_vline(xintercept = 0.08, col="red", lty="dotted", lwd=1) + 
  geom_text(aes(label=cluster),hjust=2, vjust=0)+
  scale_fill_tableau(palette = "Tableau 20") +
  theme_classic() +
  ggtitle("Unfiltered") + 
  theme(aspect.ratio = 1) 

purity.CiteFuse <- ggplot(combined[combined$method == 2,], 
                          aes(x=purity_CD4, y=purity_CD8, col=cluster)) +
  geom_point(size=4) +
  geom_hline(yintercept = 0.1, col="red", lty="dotted", lwd=1) + 
  geom_vline(xintercept = 0.08, col="red", lty="dotted", lwd=1) + 
  geom_text(aes(label=cluster),hjust=2, vjust=0)+
  scale_color_manual(values = annotation_colors$W_colors) +
  theme_classic() +
  ggtitle("CiteFuse") + 
  theme(aspect.ratio = 1) 

purity.Scrublet <- ggplot(combined[combined$method == 3,], 
                          aes(x=purity_CD4, y=purity_CD8, col=cluster)) +
  geom_point(size=4) +
  geom_hline(yintercept = 0.1, col="red", lty="dotted", lwd=1) + 
  geom_vline(xintercept = 0.08, col="red", lty="dotted", lwd=1) + 
  geom_text(aes(label=cluster),hjust=2, vjust=0)+
  scale_color_manual(values = annotation_colors$W1_colors) +
  theme_classic() +
  ggtitle("Scrublet") + 
  theme(aspect.ratio = 1) 

purity.HTODemux <- ggplot(combined[combined$method == 4,], 
                          aes(x=purity_CD4, y=purity_CD8, col=cluster)) +
  geom_point(size=4) +
  geom_hline(yintercept = 0.1, col="red", lty="dotted", lwd=1) + 
  geom_vline(xintercept = 0.08, col="red", lty="dotted", lwd=1) + 
  geom_text(aes(label=cluster),hjust=2, vjust=0) +
  scale_color_manual(values = annotation_colors$W2_colors) +
  theme_classic() +
  ggtitle("HTODemux") + 
  theme(aspect.ratio = 1) 

purity.DoubletFinder <- ggplot(combined[combined$method == 5,], 
                               aes(x=purity_CD4, y=purity_CD8, col=cluster)) +
  geom_point(size=4) +
  geom_hline(yintercept = 0.1, col="red", lty="dotted", lwd=1) + 
  geom_vline(xintercept = 0.08, col="red", lty="dotted", lwd=1) + 
  geom_text(aes(label=cluster), hjust=2, vjust=0) +
  scale_fill_tableau(palette = "Tableau 20") +
  theme_classic() +
  ggtitle("DoubletFinder") + 
  theme(aspect.ratio = 1) 

purity.combined <-  ggpubr::ggarrange(purity.unfiltered,
          purity.HTODemux,
          purity.Scrublet,
          purity.DoubletFinder,
          purity.CiteFuse,
          ncol=5, 
          legend="bottom",
          common.legend = F)
purity.combined 

purity.combined %>% ggsave(filename = "Figure4b.pdf", height=5, width = 20) 
```

#Session Info
```{r sessinf}
sessionInfo()
```