---
title: "UMAP_clustering_ouput"
author: "Hani Jieun Kim"
date: "11/03/2020"
output: 
  BiocStyle::html_document:
    code_folding: hide
    number_sections: yes
    toc: yes  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# UMAP visualistion of clustering outputs by expression matrices filtered for doublets

The script reproduces the UMAPs in Figure 4a. It is shown in Figure \@ref(fig:plot).
```{r load, echo=FALSE}
library(uwot)
library(ggplot2)
library(ggthemes)
library(viridis)
library(propr)
library(igraph)
```

Load fused expression matrices, each filtered for doublets by using different doublet detection algorithms.
```{r}
setwd("/Users/hani/Dropbox (Sydney Uni)/Single Cell Reserach/SCProj-CITEseq/codes/Code/CodeRepository")

load("fusedMatrix_CiteFuse.RData", verbose=T)
load("fusedMatrix_HTODemux.RData", verbose=T)
load("fusedMatrix_Scrublet.RData", verbose=T)
load("fusedMatrix_Unfiltered.RData", verbose=T)
load("fusedMatrix_DoubletFinder.RData", verbose=T)

load("annotation_colors.Rdata", verbose=T)
load("cluster_labels.RData", verbose = T)
```

# Generate UMAPs
```{r}
runUMAP <- function(fusedMatrix) {
    
    set.seed(2019)
  
    fusedMatrix_umap <- uwot::umap(fusedMatrix)
    
    fusedMatrix_umap_toPlot <- data.frame(
        UMAP1 = fusedMatrix_umap[, 1],
        UMAP2 = fusedMatrix_umap[, 2]
    )
    
    return(fusedMatrix_umap_toPlot)
}

#fusedMatrix_toPlot_Unfiltered <- runUMAP(fusedMatrix_Unfiltered)
#fusedMatrix_toPlot_CiteFuse <- runUMAP(fusedMatrix_CiteFuse)
#fusedMatrix_toPlot_Scrublet <- runUMAP(fusedMatrix_Scrublet)
#fusedMatrix_toPlot_HTODemux <- runUMAP(fusedMatrix_HTODemux)
#fusedMatrix_toPlot_DoubletFinder <- runUMAP(fusedMatrix_doubletFinder)

#save(fusedMatrix_toPlot_Unfiltered, 
     #fusedMatrix_toPlot_CiteFuse,
     #fusedMatrix_toPlot_Scrublet,
     #fusedMatrix_toPlot_HTODemux,
     #fusedMatrix_toPlot_DoubletFinder,
     #file="fusedMatrix_toPlot_umap0.1.4.RData")

load("fusedMatrix_toPlot_umap0.1.4.RData", verbose = T)
```
# Clustering with CiteFuse

Generate 16 clusters with spectral clustering from each similarity matrix.
```{r}
source("functions.R")

set.seed(2019)
system.time(clusters_Unfiltered <- spectralClustering(fusedMatrix_Unfiltered, K = 16, kernel = T))
system.time(clusters_Scrublet <- spectralClustering(fusedMatrix_Scrublet, K = 16))
system.time(clusters_HTODemux <- spectralClustering(fusedMatrix_HTODemux, K = 16))
system.time(clusters_DoubletFinder <- spectralClustering(fusedMatrix_doubletFinder, K = 16))
```

# Plot UMAP of fused matrix filtered by different filtering methods
```{r plot, fig.cap="UMAP of doublet filtering clustering output", fig.height=5, fig.width=20}
gg.UMAP_unfiltered <- ggplot(fusedMatrix_toPlot_Unfiltered, 
                             aes(x = UMAP1, y = UMAP2, col = as.factor(clusters_Unfiltered$labels))) +
  geom_point(alpha=0.5) +
  theme_classic() +
  scale_fill_tableau(palette = "Tableau 20") +
  theme(aspect.ratio = 1,  legend.position = "bottom") +
  labs(title = "Unfiltered")

gg.UMAP_HTODemux <- ggplot(fusedMatrix_toPlot_HTODemux, 
                           aes(x = UMAP1, y = UMAP2, col = as.factor(clusters_HTODemux$labels))) +
  geom_point(alpha=0.5) +
  theme_classic() +
  scale_color_manual(values = annotation_colors$W2_colors) +
  theme(aspect.ratio = 1,  legend.position = "bottom") +
  labs(title = "HTODemux")

gg.UMAP_Scrublet <- ggplot(fusedMatrix_toPlot_Scrublet, 
                           aes(x = UMAP1, y = UMAP2, col = as.factor(clusters_Scrublet$labels))) +
  geom_point(alpha=0.5) +
  theme_classic() +
  scale_color_manual(values = annotation_colors$W1_colors) +
  theme(aspect.ratio = 1,  legend.position = "bottom") +
  labs(title = "Scrublet")

gg.UMAP_DoubletFinder <- ggplot(fusedMatrix_toPlot_DoubletFinder, 
                                aes(x = UMAP1, y = UMAP2, col = as.factor(clusters_DoubletFinder$labels))) +
  geom_point(alpha=0.5) +
  theme_classic() +
  scale_fill_tableau(palette = "Tableau 20") +
  theme(aspect.ratio = 1,  legend.position = "bottom") +
  labs(title = "DoubletFinder")

gg.UMAP_CiteFuse <- ggplot(fusedMatrix_toPlot_CiteFuse, 
                           aes(x = UMAP1, y = UMAP2, col = as.factor(CiteFuse_cluster_labels))) +
  geom_point(alpha=0.5) +
  theme_classic() +
  scale_color_manual(values = annotation_colors$W_colors) +
  theme(aspect.ratio = 1,  legend.position = "bottom") +
  labs(title = "CiteFuse")

ggpubr::ggarrange(gg.UMAP_unfiltered,
                  gg.UMAP_HTODemux,
                  gg.UMAP_Scrublet,
                  gg.UMAP_DoubletFinder,
                  gg.UMAP_CiteFuse,
                  ncol=5,
                  common.legend = F)


ggpubr::ggarrange(gg.UMAP_unfiltered,
                  gg.UMAP_HTODemux,
                  gg.UMAP_Scrublet,
                  gg.UMAP_DoubletFinder,
                  gg.UMAP_CiteFuse,
                  ncol=5,
                  legend="bottom") %>% ggsave(filename="Figure4a.pdf", width = 20, height = 5)
```

#Session Info
```{r sessinf}
sessionInfo()
```
